---
title: "Journal (reproducible report)"
author: "Dina Abdelmguid"
date: "2023-13-05"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(knitr)
```

# Session_2 challenge:

## Task description: **SALES ANALYSIS :**

Analyze the sales by location (state) with a bar plot. Since state and city are multiple features (variables), they should be split. Which state has the highes revenue? Replace your bike_orderlines_wrangled_tbl object with the newly wrangled object (with the columns state and city).

## Solution:

**Load Libraries** 

```{r}

library(tidyverse)
library(readxl)

```

**Importing Files**

```{r}

bikes_tbl      <- read_excel("ds_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel("ds_data/01_bike_sales/01_raw_data/orderlines.xlsx")
bikeshops_tbl  <- read_excel("ds_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

```

**Joining Data**

```{r}

kable(left_join(orderlines_tbl, bikes_tbl, by = c("product.id" = "bike.id")))

```

Chaining commands with the pipe and assigning it to order_items_joined_tbl

```{r}

bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

kable(bike_orderlines_joined_tbl)
```

**Re-format:**

```{r }

bike_orderlines_joined_tbl %>% glimpse()
```

**Wrangling Data**

```{r}

bike_orderlines_joined_tbl %>% 
  select(category) %>%
  filter(str_detect(category, "^Mountain")) %>% 
  unique()

bike_orderlines_joined_tbl

bike_orderlines_wrangled_city_separated_tbl <- bike_orderlines_joined_tbl %>%
  
  #Separate category name
  
  separate(col    = category,
           into   = c("category.1", "category.2", "category.3"),
           sep    = " - ") %>% 
  
  #Separate city and state 
  
  separate(col    = location,
           into   = c("City", "State"),
           sep    = ", ") %>% 
  
  
  #Add the total price (price * quantity)  

  mutate(total.price = price * quantity)%>%
  
  #Reorganize (remove unwanted columns)
  select(-...1, -gender)%>%
  
  #Organize by pattern
  select(-ends_with(".id"))%>% 
  
  #Re-add order.id column
  
  bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>%
  
  #Re-order:
  
  select(order.id, contains("order"), contains("model"), contains("category"),
         price, quantity, total.price,
         everything()) %>%
  
  #Rename columns (one at the time vs. multiple at once)
  
  rename(bikeshop = name) %>%
  set_names(names(.) %>% str_replace_all("\\.", "_"))

kable(bike_orderlines_wrangled_city_separated_tbl, caption = "Bikes order lines separated by CITY")
```
  
**Business Insights**

```{r}

#Analyze the sales by location (state) ----

library(lubridate)
  # Step 1 - Manipulate
  sales_by_location_tbl <- bike_orderlines_wrangled_city_separated_tbl %>%
  
  # Select columns
  select(State, City, total_price) %>%
  
  # Grouping by state and summarizing sales
  group_by(State) %>% 
  summarize(state_sales = sum(total_price)) %>%
  
  # Optional: Add a column that turns the numbers into a currency format 
  # (makes it in the plot optically more appealing)
  # mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
  mutate(sales_formatted = scales::dollar(state_sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

sales_by_location_tbl
# Step 2 - Visualize
sales_by_location_tbl %>%
  
  # Setup canvas with the columns year (x-axis) and sales (y-axis)
  ggplot(aes(x = State, y = state_sales)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales_formatted)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
  # Formatting
  # scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
  # Again, we have to adjust it for euro values
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  # Rotate plot: 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  labs(
    title    = "States revenue by year",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )

#**Step 2 - Visualize**


  sales_by_location_tbl %>%
  ggplot(aes(x = State, y = state_sales)) +
  
  #Geometries
  
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales_formatted)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
  #Formatting
  scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
  #Again, we have to adjust it for euro values
  
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  #Rotate plot: 
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  labs(
    title    = "States revenue by year",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )
```

**Analyze the sales by location and year**

```{r}

#Step 1 - Manipulate
sales_by_state_year_tbl <- bike_orderlines_wrangled_city_separated_tbl %>%
  
  # Select columns and add a year
  select(order_date, total_price, State) %>%
  mutate(year = year(order_date)) %>%
  
  # Group by and summarize year and main catgegory
  group_by(State, year) %>%
  summarise(sales = sum(total_price)) %>%
  ungroup() %>%
  
  # Format $ Text
  mutate(sales_formatted = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))
```


```{r plot, fig.width=10, fig.height=7}

# Step 2 - Visualize
sales_by_state_year_tbl %>%
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = State)) +
  
  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot
  
  # Facet
  facet_wrap(~ State) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title = "Revenue by year and state",
    subtitle = "Each State has a dynamic trend",
    fill = "States" # Changes the legend name
  )

```



<<<<<<< HEAD
=======



>>>>>>> parent of 8755b95 (journal_2)
# Session_5 Data visualization:

## Task description

Challenge 1
Goal: Map the time course of the cumulative Covid-19 cases!

Challenge 2
Goal: Visualize the distribution of the mortality rate (deaths / population)


## Solustion

```{r}
#Import required Libraries

library(scales)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)
library(ggthemes)
library(dplyr)
library(maps)
```

**Task 1** 

```{r}

covid_data_tbl <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")

#Table for Challenge 1 before plot

  covid_data_select_tbl<- covid_data_tbl %>%
  select(countriesAndTerritories,cases,dateRep,month,year,day)%>%
  relocate(year,month,day)%>%
  filter(year==2020,month>1) %>%
  filter(day!=1)%>%
  filter(countriesAndTerritories=="France"|countriesAndTerritories=="Germany"|countriesAndTerritories=="United_Kingdom"|countriesAndTerritories=="Spain"|countriesAndTerritories=="United_States_of_America")%>%
  group_by(countriesAndTerritories,month)%>%
  summarize(totalcases = sum(cases)) %>%
  ungroup()
    
kable(covid_data_select_tbl, caption = "Filtered data")  

```

**Prepared Plots**

```{r}
#Prepare plot
  covid_data_select_tbl%>%
  ggplot(aes(month ,totalcases, color = countriesAndTerritories)) +
        geom_smooth(method = "loess", span = 0.2)+
        scale_y_continuous(labels = scales::dollar_format(scale  = 1/1e6, 
                                                        prefix = "", 
                                                        suffix = "M")) +
        scale_x_continuous(breaks = seq(2, 11 , by=1),labels= c("February","March","April","May","June","July","August","September","October","November")) +
 # scale_x_continuous(labels = scales::dollar_format(scale = 1/1e6,
                                                     #prefix= "",
                                                    # suffix= "February")) +
   
labs(
  title = ("Covid-19 confirmed cases worldwide"),
  subtitle = ("United States has the highest rate of cases"),
  caption = "",
  x = "(Year 2020)",
  y = "Cumulative Cases",
  color = "Country"

          )+
    geom_label(aes(label = (totalcases)), 
              hjust = "inward",
              size  = 3,
              color = RColorBrewer::brewer.pal(n = 11, name = "RdBu")[11]) 
    
```

**Task 2**

\#World data table:

```{r}
#importing data

covid_data_tbl <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")

world <- map_data("world")%>%mutate(across(region, str_replace_all, "_", " ")) %>%
  mutate(region = case_when(
    
    region == "UK"~ "United_Kingdom",
    region == "USA"~"United_States_of_America" ,
    region == "Czech_Republic"~"Czechia",
    TRUE ~ region
    
  ))
covid_data_tbl%>%mutate(across(countriesAndTerritories, str_replace_all, "_", " ")) %>%
  mutate(countriesAndTerritories = case_when(
    
    countriesAndTerritories == "United Kingdom" ~ "United_Kingdom",
    countriesAndTerritories == "United_States_of_America" ~ "United States of America",
    countriesAndTerritories == "Czechia"~"Czechia",
    TRUE ~ countriesAndTerritories
    
  ))

#manipulation of world data table
world_map<-world%>%select(region,long,lat,group)%>%rename(countriesAndTerritories=region)

```

\#Covid data:

```{r}
#manipulation of covid data table
covid_modified_data_tbl<- covid_data_tbl%>%select(day,month,year,countriesAndTerritories,deaths,popData2019)%>%
  group_by(year,countriesAndTerritories,popData2019)%>%
  summarise(total_death=sum(deaths))%>%
  ungroup()%>%
  mutate(mortality_rate=(total_death/popData2019)*100)

#merging data between 2 tables 
All_data_tbl<-left_join(covid_modified_data_tbl,world_map,by="countriesAndTerritories")%>%filter(year==2020)
```

**Prepared Plots**

```{r}

#first layer of the map
world_map <- map_data("world")
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightgray", colour = "black",size=0.1)

#second layer of the map
ggplot(data=All_data_tbl, aes(x=long, y=lat, group = group))+
  geom_polygon(aes(fill = mortality_rate), color = "red",size=0.1)+
  scale_fill_viridis_c(option = "C", alpha = 0.75 )
```



